An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-02 rule-ltl.tex
fixing problems

The order of calculating $\current$ is subtle here, because
auxiliary variables can appear both on the left and right of their
definition. For PLTL, one can start calculating the subformulas
bottom up or top down. But for EPLTL, this would not work. To see this, consider, for example, the formula~(\ref{form3}). It contains
the definition $q := \ominus \neg q$. Now, we cannot calculate this
bottom up, since the left-hand $q$ is not calculated yet.
If we calculate it top down, we will encounter the same problem.
However, notice that the top down calculation does not need the
value of $q$ to calculate $\ominus \neg q$; this value should already be available as $\old ( \neg q )$.

Under {\em mixed evaluation order}, one calculates $\current$ in the following order. The calculation within each of the four categories can be performed top down or bottom up.
\begin{itemize}
\item Calculate 
$\current ( \delta )$ for each subformula $\delta$
that appears in $\varphi_j$ of rule
$a_j := \varphi_j$ but not within the scope of
a $\ominus$ operator. (This is not
a problem since for $\delta = \ominus \gamma$,
$\current ( \delta )$ is set to $\old ( \gamma )$.)
\item Set 
$\current ( a_j )$ to $\current ( \varphi_j )$ for each $j$.
\item Calculate $\current ( \delta )$ for each subformulas $\delta$
that appear in $\varphi_j$ of a rule
$a_j = \varphi_j$ {\em  within} the scope of
a $\ominus$ operator.
\item Calculate $now ( \psi )$ (using $\current ( a_j )$) for
the statement $\psi$.
\end{itemize}

@@

Assume a finite set of infinite domains\footnote{Finite domains are handled with some minor changes, see~\cite{HPU}.}
$D_1, D_2, \ldots$,
e.g., integers or strings. 
Let $V$ be a finite set of {\em variables}, 
with typical instances $x$, $y$, $z$.
%We denote by $x: D$ the fact that the variable $x$ has the domain $D$.
An {\em assignment} over a set of variables $V$
maps each variable $x \in V$ to a value from
its associated domain $\domain ( x )$, where multiple variables (or all of them)
can be related to the same domain. For example
$[ x \rightarrow 5 , y \rightarrow \text{``abc''} ]$ assigns
the values $5$ to $x$ and the value ``abc'' to $y$.
Let $T$ a set of {\em relation} 
with typical instances $R$.
Each relation is associated with some position-dependent domains.

We defined models for FLTL based on {\em temporal} relations~\cite{Chomicki}. That is,
their last parameter
is a natural number. For a relation $R$, $R [ i ]$ is the relation obtained from $R$ by
restricting it to the value $i$ in the last parameter. For simplicity, we will describe henceforth
the logic with relations $R$ that have exactly two parameters
, hence $R [ i ]$ will have just one parameter whose domain will be denoted as $d ( R )$; this domain is fixed over all the value of the second parameter $i$. The the definition of the logic over other number of parameters is quite straightforward, and our implementation of runtime verification fully supports it.

>>>>

The order of calculating $\current$ is subtle here, because
auxiliary variables can appear recursively on the right-hand 
side of their definition. For PLTL, one can start calculating the subformulas
bottom up or top down. But for EPLTL, this would not work. To see this, consider, for example, the formula~(\ref{form3}). It contains
the definition $q := \ominus \neg q$. Now, we cannot calculate this
bottom up, since the left-hand $q$ is not calculated yet, and we need its value to compute the right-hand side $q$ leaf node.
If we calculate it top down, we will encounter the same problem.
However, notice that the top down calculation does not need the
value of $q$ to calculate $\ominus \neg q$; this value should already be available as $\old ( \neg q )$.

\klaus{I don't see that we need to talk about the top-down evaluation.}

Under {\em mixed evaluation order}, one calculates $\current$ in the following order. The calculation within each of the four categories can be performed top down or bottom up.

\begin{enumerate}
\item Calculate 
$\current ( \delta )$ for each subformula $\delta$
that appears in $\varphi_j$ of rule
$a_j := \varphi_j$ but not within the scope of
a $\ominus$ operator. (This is not
a problem since for $\delta = \ominus \gamma$,
$\current ( \delta )$ is set to $\old ( \gamma )$.)
\item Set 
$\current ( a_j )$ to $\current ( \varphi_j )$ for each $j$.
\item Calculate $\current ( \delta )$ for each subformula $\delta$
that appears in $\varphi_j$ of a rule
$a_j = \varphi_j$ {\em  within} the scope of
a $\ominus$ operator.
\item Calculate $\current ( \delta )$ for each subformula $\delta$
that appears in the statement $\psi$ - using $\current ( a_j )$.
\end{enumerate}

@@

Assume a finite set of infinite domains\footnote{Finite domains are handled with some minor changes, see~\cite{HPU}.}
$D_1, D_2, \ldots$,
e.g., integers or strings. 
Let $V$ be a finite set of {\em variables}, 
with typical instances $x$, $y$, $z$.
%We denote by $x: D$ the fact that the variable $x$ has the domain $D$.
An {\em assignment} over a set of variables $V$
maps each variable $x \in V$ to a value from
its associated domain $\domain ( x )$, where multiple variables (or all of them)
can be related to the same domain. For example
$[ x \rightarrow 5 , y \rightarrow \text{``abc''} ]$ assigns
the values $5$ to $x$ and the value ``abc'' to $y$.
Let $T$ be a set of {\em relations} 
with typical instances $R$.
Each relation is associated with some position-dependent domains.

\klaus{Above: let T be ... Is T used later? The relations are used in the next paragraph, why not introduce them there? Also, relations are not defined really.}

\klaus{I don't quite understand the motivation below for this way of modeling events: relations with a time stamp, and the whole story of the time stamp not being mentioned etc. Why not just the usual models: traces? Too late to change now for CAV though.}

We define models for FLTL based on {\em temporal} relations~\cite{Chomicki}. That is,
their last parameter
is a natural number. For a relation $R$, $R [ i ]$ is the relation obtained from $R$ by
restricting it to the value $i$ in the last parameter. For simplicity, we will describe henceforth
the logic with relations $R$ that have exactly two parameters
, hence $R [ i ]$ will have just one parameter whose domain will be denoted as $d ( R )$; this domain is fixed over all the value of the second parameter $i$. The the definition of the logic over other number of parameters is quite straightforward, and our implementation of runtime verification fully supports it.
